---
title: "Fetching FHRS data the easy way"
author: "Helen Graham"
date: "24 May 2019"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fetching FHRS data the easy way}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

### Why make an R package for FHRS data?

The FHRS is currently published as open data, in the form of individual XML files 
for each local authority, or available through our API. Neither of those are a 
totally straightforward way to get the data you want. 

Enter `fhrsdata`. This package wraps useful API calls in R functions, which can 
be customised via the functions' parameters, to provide data extracts for a 
range of scenarios. 

### Installing the package

```{r warning=FALSE, message=FALSE}
#install devtools if necessary
#install.packages("devtools")
devtools::install_github("helen-food/fhrsdata", quiet = TRUE)
```

### Load it up

```{r warning=FALSE, message=FALSE}
library(fhrsdata)
```


### Fetching establishments

At the heart of the package is the `get_establishments()` function. This calls 
upon the API to return a tibble of establishments from the FHRS. You could 
call this function with no arguments and it would return all the establishments 
in the data, although it would take a very long time, as there are over half 
a million of them. Not recommended!

The chances are, you do not want everything in the FHRS. You may only want places 
people go to eat (restaurants, cafes, mobile caterers, etc.) and not all the 
schools and hospitals. Maybe you are only interested in retailers, or a 
particular retailer. Maybe you only want to know about a particular place. 
Maybe you only want to know how many pubs there are called "The Red Lion". All of 
these are possible with the `get_establishments()` function.

The `name` argument (NULL by default) is used to search for establishments by name. 
For example: Show me all the establishments with "red lion" in the name:

```{r}
get_establishments(name = "red lion")
```

Almost 600 red lions! However, are these all pubs?

```{r warning=FALSE, message=FALSE}
library(dplyr)
get_establishments("red lion") %>% count(BusinessType)
```


Some are not. What if you only want pubs? You would use the `type` argument. But 
first, you need to know the code for the type you want to select. You find this 
out using the `get_business_types()` function.

```{r}
get_business_types()
```

From this we can see that the code for pubs is 7843. So let's add that to our 
API call:

```{r}
get_establishments(name = "red lion", type = 7843)
```

Now only the pubs have been returned. The final customisation option is to filter 
the search by local authority. Out of interest, which local authorities have the 
most Red Lions?

```{r}
get_establishments(name = "red lion", type = 7843) %>% 
  count(LocalAuthorityName) %>% 
  arrange(desc(n)) %>% 
  head()
```

And the winner is Powys. Let's say we wanted to further restrict our search for 
red lions just to Powys for some reason. First we need to find out the ID number 
for Powys in the data. We can do that using the `get_authorities()` function, which 
returns, you've guessed it, a list of all local authorities in the FHRS. That list 
is a bit longer than the list of business types, though. Fortunately we know that 
Powys is in Wales, and can search only for authorities in Wales using the `country` 
argument.

```{r}
get_authorities(country = "Wales")
```

Or perhaps even better, just filter the authority search down to Powys.

```{r}
get_authorities() %>% 
  filter(Name == "Powys")
```


To complete our slightly implausible quest for Red Lion pubs in Powys, the 
`get_establishments()` function is called as follows:

```{r}
get_establishments(name="red lion", type = 7843, la = 349)
```

Finally, it is worth noting that the tibble returned by default from `get_establishments()` is 
a condensed version that includes only a selection of the (arguably most useful) fields. But 
if you want to make that decision for yourself, you can return everything by 
setting the `condensed` argument to FALSE.

```{r}
get_establishments(name="red lion", type = 7843, la = 349, condensed = FALSE)
```

